Palmer Archipelago (Antarctica) penguins: an alternative to Edgar Anderson''s

### 2020-07-11 manuscript/manuscript_text.Rmd
add tidy kmeans

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(palmerpenguins)
library(kableExtra)
library(patchwork)
library(gt)
library(GGally)
library(recipes)
library(shadowtext)
knitr::opts_chunk$set(echo = TRUE)

@@

p_bills <- penguins %>% 
  select(species, bill_length_mm, bill_depth_mm) %>% 
  drop_na()

p_bills_nospecies <- p_bills %>% 
  select(-species) %>% 
  scale() %>% 
  as.data.frame()

# Perform k-means on penguin bill dimensions (k = 3, w/20 centroid starts)
p_bills_kmeans <- kmeans(p_bills_nospecies, centers = 3, nstart = 20)

# Bind clusters to the data
p_kmeans_clusters <- cbind(p_bills_nospecies, cluster = as.factor(p_bills_kmeans$cluster), species = p_bills$species)

# Plot clusters
p_kmeans_gg <- ggplot(p_kmeans_clusters, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_text(aes(label = cluster, color = species)) +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4"))

# Get counts in each cluster by species
p_cluster_counts <- p_kmeans_clusters %>% 
  count(species, cluster) %>% 
  pivot_wider(names_from = species, values_from = n) %>% 
  arrange(cluster)

@@

iris_petals <- iris %>% 
  select(Species, Petal.Length, Petal.Width)

# remove species factor & scale petal dimensions: 
ip_nospecies <- iris_petals %>% 
  select(-Species) %>% 
  scale() %>% 
  as.data.frame()

# Perform k-means on iris petal dimensions (k = 3, w/20 centroid starts)
set.seed(1967)
ip_kmeans <- kmeans(ip_nospecies, centers = 3, nstart = 20)

# Bind clusters to the data
iris_kmeans_clusters <- cbind(ip_nospecies, cluster = as.factor(ip_kmeans$cluster), species = iris$Species)

# Plot clusters: 
iris_kmeans_gg <- ggplot(iris_kmeans_clusters, aes(x = Petal.Length, y = Petal.Width)) +
  geom_text(aes(label = cluster, color = species)) +
   theme(legend.position = "bottom",
         panel.background = element_rect(fill = "gray92", color = "gray90")) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4"))

# Get counts in each cluster by species
iris_cluster_counts <- iris_kmeans_clusters %>% 
  count(species, cluster) %>% 
  pivot_wider(names_from = species, values_from = n) %>% 
  arrange(cluster)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
(p_kmeans_gg | iris_kmeans_gg) + plot_annotation(tag_levels = "A")

kmeans_2var_table <- cbind(p_cluster_counts, iris_cluster_counts) %>% 
  kable() %>% 
  kable_styling(full_width = FALSE) %>% 
  column_spec(c(5:8), background = "gainsboro") %>% 
  add_header_above(c("Penguins cluster assignments" = 4, "Iris cluster assignments" = 4))

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(palmerpenguins)
library(kableExtra)
library(patchwork)
library(gt)
library(GGally)
library(recipes)
library(shadowtext)
library(broom)
knitr::opts_chunk$set(echo = TRUE)

@@

pb_species <- penguins %>% 
  select(species, starts_with("bill")) %>% 
  drop_na()

pb_nospecies <- pb_species %>% 
  select(-species) %>% 
  recipe() %>% 
  step_normalize(all_numeric()) %>% 
  prep() %>% 
  juice()

# Perform k-means on penguin bill dimensions (k = 3, w/20 centroid starts)
# save augmented data
pb_clust <- 
  pb_nospecies %>% 
  kmeans(centers = 3, nstart = 20) %>% 
  broom::augment(pb_species)

# Plot clusters
pb_kmeans_gg <- 
  pb_clust %>% 
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_text(aes(label = .cluster, color = species),
            key_glyph = draw_key_rect) +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4"))

# Get counts in each cluster by species
pb_clust_n <- pb_clust %>% 
  count(species, .cluster) %>% 
  pivot_wider(names_from = species, values_from = n, names_sort = TRUE) %>% 
  arrange(.cluster)
```

@@

ip_species <- iris %>% 
  select(Species, starts_with("Petal"))

# remove species factor & scale petal dimensions: 
ip_nospecies <- ip_species %>% 
  select(-Species) %>% 
  recipe() %>% 
  step_normalize(all_numeric()) %>% 
  prep() %>% 
  juice()

# Perform k-means on iris petal dimensions (k = 3, w/20 centroid starts)
ip_clust <- 
  ip_nospecies %>% 
  kmeans(centers = 3, nstart = 20) %>% 
  broom::augment(ip_species)

# Plot clusters
ip_kmeans_gg <- 
  ip_clust %>% 
  ggplot(aes(x = Petal.Length, y = Petal.Width)) +
  geom_text(aes(label = .cluster, color = Species), 
            key_glyph = draw_key_rect) +
  theme(legend.position = "bottom",
         panel.background = element_rect(fill = "gray92", color = "gray90")) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4"))

# Get counts in each cluster by species
ip_clust_n <- ip_clust %>% 
  count(Species, .cluster) %>% 
  pivot_wider(names_from = Species, values_from = n, names_sort = TRUE) %>% 
  arrange(.cluster)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
(pb_kmeans_gg | ip_kmeans_gg) + plot_annotation(tag_levels = "A")

kmeans_2var_table <- cbind(pb_clust_n, ip_clust_n) %>% 
  kable() %>% 
  kable_styling(full_width = FALSE) %>% 
  column_spec(c(5:8), background = "gainsboro") %>% 
  add_header_above(c("Penguins cluster assignments" = 4, "Iris cluster assignments" = 4))
